Cropland Bias and Accuracy

### 2015-10-18 Rmd/cropland-bias.Rmd
18/10/2015: Updates through yield/production analysis. See changelog for specifics.

## Plots
Plot maps of biases at different resolutions, for publication output (and display here). To avoid the visual distortion of NA areas filled in by larger levels of aggregation, each aggregated difference must be disaggregated first and then masked before plotting.   
```{r, eval = FALSE}
namask <- raster(full_path(p_data, "namask.tif"))
# cellStats(namask, sum)
snms <- c("sa30", "globmu", "modmu", "geow")  # all lc data (no MOD/GC extremes)
lev <- c("f1", "f25", "f50", "f100")  # just show four levels

@@

# Total bias

Calculate how much cropland estimates differ between datasets
```{r, eval = FALSE}
paths <- full_path(p_data, dir(p_data, paste0("cover*.*sum_mask.tif$")))
gti <- lapply(paths, raster)
names(gti) <- c("g2007", "g2011")
gti <- lapply(gti, function(x) mask(x, namask))  # apply namask
snms <- c("sa30", "globmu", "modmu", "geow")  # all lc data (no MOD/GC extremes)

@@

In relation to calculation method (across country, agricultural region only, or bin-wise (cropland density classes) mean.
```{r, eval = FALSE}
load(full_path(p_data, "lcf_out2011.rda"))
lcf <- data.table(lcf_out11)

@@

pdf(full_path(p_fig, "cropland_bias_scale.pdf"), height = 7, width = 7)
par(mar = rep(1, 4), oma = c(2, 2, 0, 0), mgp = c(1, 0.5, 0), tcl = -0.3)
plot(xl, yl, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
     ylab = "", xlab = "")#,
for(i in shd) polyfunc2(x = i, y = yl, w = 3, col = g1, bcol = g1, lwd = 1)
abline(h = yax, v = NULL, col = "grey80", lty = 1)
polyfunc2(x = 8.75, y = yl, w = 18.5, col = "transparent", bcol = "black")
lines(c(-1, 18), c(0, 0), lwd = 2, col = "grey80")
# i <- 1; j <- 1; k <- 1
for(i in 1:length(svec)) {
  lv <- colnms[i]
  for(j in 1:length(snms)) {
    pchs1 <- pchs
    nm <- lcnms[j]
    for(k in 1:length(mutype)) {
      #dat <- get(mutype[k])[V1 == snms[j] & V2 %in% LC2, get(lev_vec[i])]
      dat <- lcf[Metric == mutype[k] & Map == nm, lv, with = FALSE]
      pcol <- makeTransparent(cols[j], alpha = alph[k])
      polyfunc2(xa[j, i] + o[k], range(dat), w = w, col = pcol, bcol = pcol)
      for(z in 1:length(reg)) {
        lpfunc(x[i] + xo[j] + o[k], dat[z], col = "black", size = cx[z], 
               pch = pchs[z], type = "pt")
      }
    }
  }
}

@@

>>>>

## Plots
Plot maps of biases at different resolutions, for publication output (and display here). To avoid the visual distortion of NA areas filled in by larger levels of aggregation, each aggregated difference must be disaggregated first and masked before plotting.   
```{r, eval = FALSE}
namask <- raster(full_path(p_data, "namask.tif"))
# cellStats(namask, sum)
snms <- c("sa30", "globmu", "modmu", "geow")  # all lc data (no MOD/GC extremes)
lev <- c("f1", "f25", "f50", "f100")  # just show four levels

@@

# Total bias
Calculate how much cropland estimates differ between datasets
```{r, eval = FALSE}
paths <- full_path(p_data, dir(p_data, paste0("cover*.*sum_mask.tif$")))
gti <- lapply(paths, raster)
names(gti) <- c("g2007", "g2011")
gti <- lapply(gti, function(x) mask(x, namask))  # apply namask
snms <- c("sa30", "globmu", "modmu", "geow")  # all lc data (no MOD/GC extremes)

@@

## Primary method

Weighted by __actual__ cropland percentage, and then, with aggregation, by number of pixels being aggregated.

```{r, eval = FALSE}
# Create weights maks
fact <- c(5, 10, 25, 50, 100)
namask <- raster(full_path(p_data, "namask.tif"))  # NA mask
# cellStats(!is.na(gti$g2011), sum)
# cellStats(namask, sum)

namask2 <- !is.na(namask)
awgts <- aggregate_rast_list(fact, list(namask2), fun = sum)
awgts <- lapply(awgts, function(x) x[[1]])

# Subset difference grids, for 2011
snms <- c("sa30", "globmu", "modmu", "geow")
err2011 <- lapply(lev, function(x) {
  sapply(dlist_act[snms], function(y) list(y[[x]]$g2011))
})
names(err2011) <- lev

gti_agg <- aggregate_rast_list(fact, gti)  # GTI rasters
gti_agg11 <- lapply(gti_agg, function(x) x$g2011)

# calculate bias/MAE for data
a <- bias_statsw_list(gti_agg11, awgts, err2011, snms, wm, "mu")
b <- bias_statsw_list(gti_agg11, awgts, err2011, snms, wma, "mua")
bacc2011 <- rbind(a, b)

# check error stats - do they match alternate approaches?
for(i in c("f1", "f25", "f10")) {
  for(j in c("sa30", "modmu", "geow")) { 
    print(paste("cross-checking calculations in", i, j))
    a1 <- bias_statsw(gti_agg11[[i]], awgts[[i]], err2011[[i]], snms, wm, 
                      "mu", aweight = FALSE)[, j, with = FALSE]
    a2 <- bias_statsw(gti_agg11[[i]], awgts[[i]], err2011[[i]], snms, wm, 
                      "mu", rweight = FALSE, aweight = FALSE)[,j, with=FALSE]
    a3 <- bias_statsw(gti_agg11[[i]], awgts[[i]], err2011[[i]], snms, wm, 
                      "mu")[, j, with = FALSE]
    c1 <- getValues(err2011[[i]][[j]])
    w1 <- getValues(gti_agg11[[i]])
    w2 <- getValues(awgts[[i]])
    print("non-area weighted mean matches?")
    print(a1 == round(weighted.mean(c1, w1, na.rm = TRUE), 2))
    print("totally unweighted mean matches?")
    print(a2 == round(cellStats(err2011[[i]][[j]], mean), 2))
    print("double weighted mean matches?")
    print(a3 == round(weighted.mean(c1, w1 * w2, na.rm = TRUE), 2))
  }
}

# Subset difference grids, for 2007
snms <- c("sa30", "globmu", "modmu", "geow")
err2007 <- lapply(lev, function(x) {
  sapply(dlist_act[snms], function(y) list(y[[x]]$g2007))
})
names(err2007) <- lev
gti_agg07 <- lapply(gti_agg, function(x) x$g2007)

a <- bias_statsw_list(gti_agg07, awgts, err2007, snms, wm, "mu")
b <- bias_statsw_list(gti_agg07, awgts, err2007, snms, wma, "mua")
bacc2007 <- rbind(a, b)

fact <- c(1, 5, 10, 25, 50, 100)
alph <- c(225, 40)
x <- c(0, 3, 6, 9, 12, 15)
w <- 3 / 8
xo <- (cumsum(rep(w, 8)) - w / 2)[-c(2, 4, 6, 8)]
o <- c(0, w)
xa <- sapply(x, function(x) x + xo)
cx <- c(1.25, 1, 1)
g1 <- "grey90"
reg <- c("Country", "Agricultural", "Density independent")
lcnms <- c("SA-LC", "GlobCover", "MODIS", "GeoWiki")
mutype <- c("mua", "mu")
# svec <- paste(gsub("f", "", bacc2011[stnm == "mu", ol]), "km")
svec <- bacc2011[stnm == "mu", ol]
xl <- c(-0.5, 18)
yl <- c(-30, 50)
shd <- c(4.5, 10.5, 16.5)
cols <- c("red", "orange3", "green4", "blue")
colnms <- snms
pchs <- c("+", "o", "x")
yax <- seq(yl[1], yl[2], 5)
xax <- seq(1, 6, 5 / (length(yax) - 1))

pdf(full_path(p_fig, "cropland_bias_main.pdf"), height = 7, width = 7)
par(mar = rep(1, 4), oma = c(2, 2, 0, 0), mgp = c(1, 0.5, 0), tcl = -0.3)
plot(xl, yl, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
     ylab = "", xlab = "")#,
for(i in shd) polyfunc2(x = i, y = yl, w = 3, col = g1, bcol = g1, lwd = 1)
abline(h = yax, v = NULL, col = "grey80", lty = 1)
polyfunc2(x = 8.75, y = yl, w = 18.5, col = "transparent", bcol = "black")
lines(c(-1, 18), c(0, 0), lwd = 2, col = "grey80")
# i <- 1; j <- 1; k <- 1
for(i in 1:length(svec)) {
  lv <- svec[i]
  for(j in 1:length(snms)) {
    # pchs1 <- pchs
    nm <- snms[j]
    for(k in 1:length(mutype)) {
      dat <- bacc2011[ol == lv & stnm == mutype[k], nm, with = FALSE][[1]]
      pcol <- makeTransparent(cols[j], alpha = alph[k])
      polyfunc2(xa[j, i] + o[k], c(0, dat), w = w, col = pcol, bcol = pcol)
    }
  }
}
axis(1, at = seq(1.5, 18.5, 3), labels = fact)
axis(2, at = yax, labels = yax, las = 2)
mtext(side = 1, text = "Resolution (km)", outer = TRUE, line = 0.5)
mtext(side = 2, text = "Bias/MAE (%)", outer = TRUE, line = 1)
legend(x = 12.4, y = -19.9, legend = lcnms, pch = 15, col = cols, adj = 0,
       pt.cex = 1.5, bty = "n", cex = 0.8, x.intersp = 0.5)
legend(x = 11.9, y = -19.9, legend = rep("", 4), pch = 15, adj = 0, 
       col = makeTransparent(cols, alpha = alph[2]), pt.cex = 1.5, bty = "n", 
       cex = 0.8)
text(x = 12.7, y = -20.5, labels = "MAE", srt = 45, adj = c(0, 0), cex= 0.8)
text(x = 12.2, y = -20.5, labels = "Bias", srt = 45, adj = c(0, 0), cex = 0.8)
dev.off()

## Other calculations methods

Across country, agricultural region only, or bin-wise (cropland density classes) mean.
```{r, eval = FALSE}
load(full_path(p_data, "lcf_out2011.rda"))
lcf <- data.table(lcf_out11)

@@

pdf(full_path(p_fig, "cropland_bias_region.pdf"), height = 7, width = 7)
par(mar = rep(1, 4), oma = c(2, 2, 0, 0), mgp = c(1, 0.5, 0), tcl = -0.3)
plot(xl, yl, pch = "", yaxt = "n", xaxt = "n", xaxs = "i", yaxs = "i", 
     ylab = "", xlab = "")#,
for(i in shd) polyfunc2(x = i, y = yl, w = 3, col = g1, bcol = g1, lwd = 1)
abline(h = yax, v = NULL, col = "grey80", lty = 1)
polyfunc2(x = 8.75, y = yl, w = 18.5, col = "transparent", bcol = "black")
lines(c(-1, 18), c(0, 0), lwd = 2, col = "grey80")
# i <- 1; j <- 1; k <- 1
for(i in 1:length(svec)) {
  lv <- colnms[i]
  for(j in 1:length(snms)) {
    pchs1 <- pchs
    nm <- lcnms[j]
    for(k in 1:length(mutype)) {
      #dat <- get(mutype[k])[V1 == snms[j] & V2 %in% LC2, get(lev_vec[i])]
      dat <- lcf[Metric == mutype[k] & Map == nm, lv, with = FALSE]
      pcol <- makeTransparent(cols[j], alpha = alph[k])
      polyfunc2(xa[j, i] + o[k], range(dat), w = w, col = pcol, bcol = pcol)
      for(z in 1:length(reg)) {
        lpfunc(x[i] + xo[j] + o[k], dat[z], col = "black", size = cx[z], 
               pch = pchs[z], type = "pt")
      }
    }
  }
}

@@

# check bin frequencies starting a bit higher: 2.5% cover
binv <- seq(2.5, 97.5, 5)
gti_bins <- lapply(gti_agg, function(x) {
  lapply(x, function(y) cut(y, breaks = binv, include.lowest = TRUE))
})
bins_25 <- lapply(gti_bins, function(x) freq(x$g2011))
barplot(bins_25$f1[1:19, 2])
```
